---
title: "Covid91 vaccine study Final2022"
author: "Rabi Elkhoury"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final_Data.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
options(digits = 2)
```


# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

```{r}
Males <- filter(FinalData, sex == "M")
Females <- filter(FinalData, sex == "F")
LGBTQ <- filter(FinalData, LGBTQ == "gay")
DrugUsers <- filter(FinalData, DrugUser == "yes")
```



## Males


### Descriptive Results

#### Graphical Results
```{r}
dd2 <- Males %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

Generating a barplot via ggplot for the data involving the incidence of MALES with COVID-19, along with whether or not they were treated with the vaccine, shows that, though a higher frequency of persons didn't get COVID-19 at all, most of those who did had taken the placebo.  Around 67% of males who contracted COVID-19 had taken the placebo, compared to only about 33% of males who had gotten the disease.  However, more testing is required to ensure such results are statistically significant.  


#### Numerical Results
```{r}
table1 <- xtabs(~treatment + infected, data=Males)
rowPerc(table1)
colPerc(table1)
```
Our numerical results corroborate the graphical descriptive results created above; we can see that, among Males who got COVID-19, 68% of them had taken the placebo, as compared to 32% of males who had gotten the vaccine.  While a majority of males didn't get COVID-19 at all, 3% of those men who received the placebo got the virus, as compared to only 1% of males who got the vaccine.  This further supports the notion that the vaccine aids in protection against the vaccine, but inferential results are required. 



### Inferential Results
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
malefisher <- fisher.test(table1)
```
Multiple data points generated from our Chi-Square and Fisher Exact tests lend credibility to our earlier tests.  The observed counts from the Chi-Square test show convey that over 100 more males who got the placebo ended up getting COVID-19, as compared to their vaccinated fellows.  Our odds ratio is 2.1, implying that males who get placebo-based treatment are 2.1 times as likely to get COVID-19 as compared to their vaccinated counterparts.  Furthermore, our p-value is infinitesimal, and such a small value suggests our results are statistically significant.  





### Efficacy of Vaccine for Males
```{r}
MaleEfficacy <- malefisher$estimate/(1 + malefisher$estimate) * 100
malefisherlow <- malefisher$conf.int[1]
malefisherhigh <- malefisher$conf.int[2]
EfficacyMalesLow <- ((malefisherlow)/(1 + malefisherlow) * 100)
EfficacyMalesHigh <- ((malefisherhigh)/(1 + malefisherhigh) * 100)
```
The estimated efficacy of the vaccine among male patients is `r MaleEfficacy`%.  The 95% confidence interval for the efficacy of the vaccine pertaining to males is `r EfficacyMalesLow`% to `r EfficacyMalesHigh`%.




## Females

### Descriptive Results

#### Graphical Results

```{r}
dd3 <- Females %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicG <- ggplot(dd3,aes(x=infected,y=count,fill=treatment))
basicG + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicGG <- ggplot(dd3,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicGG + geom_bar(stat="identity", position = "dodge")
```

Analyzing the bar-plots pertaining to the female COVID-19 infection rate, along with their vaccination status (having received a bonafide vaccine or the placebo) yields results that somewhat mirror those of the male cases.  Our frequency-based bar-plot exhibits that, although most participants did not contract COVID-19 at all, a larger count of those who did seemed to have taken the placebo.  Shifting focus to the proportion-based bar=plot, a strong majority (nearly 60%) of COVID-19 positive cases had taken the placebo, while a lesser minority had taken the vaccine.  Additional numerical and inferential results are necessary to come to a final conclusion on these results.  

#### Numerical Results

```{r}
table2 <- xtabs(~treatment + infected, data=Females)
rowPerc(table2)
colPerc(table2)
```
Viewing our data from a tabular lens further justifies the earlier graphical results investigated; clearly, an immense majority of female persons didn't get COVID-19 at all (note: data displays in whole numbers because of options(digits = 1) command), but a majority of those who did (59%) had taken the placebo instead of the vaccine (41%).  As with the male data, there is mounting evidence that those who received the vaccine may be notably protected against COVID-19 infection, but legitimate inferential test is required to validate this.  


### Inferential Results

```{r}
chisq.test(table2)
chisqtestGC(table2)
fisher.test(table2)
femalefisher <- fisher.test(table2)
```
Running our Chi-Square and FisherExact tests confirm our earlier established assumptions about the female COVID-19 statistics collected.  Our observed counts indicate that 64 more females who took the placebo contracted COVID-19, as compared to females who genuinely been inoculated.  Our odds ratio was calculated to be 1.4; this can be interpreted as that those females who only took the placebo have a 1.4 times higher chance of contracting COVID-19, compared to their vaccinated counterparts.  Furthermore, our p-value computation is essentially zero, affirming that our results are statistically significant.  


### Efficacy of Vaccine for Females
```{r}
FemaleEfficacy <- femalefisher$estimate/(1 + femalefisher$estimate) * 100
Femalefisherlow <- femalefisher$conf.int[1]
Femalefisherhigh <- femalefisher$conf.int[2]
EfficacyFemalesLow <- ((Femalefisherlow)/(1 + Femalefisherlow) * 100)
EfficacyFemalesHigh <- ((Femalefisherhigh)/(1 + Femalefisherhigh) * 100)
```
The estimated efficacy of the vaccine among male patients is `r FemaleEfficacy`%.  The 95% confidence interval for the efficacy of the vaccine pertaining to females is `r EfficacyFemalesLow`% to `r EfficacyFemalesHigh`%.


## LGBTQ

### Descriptive Results


#### Graphical Results
```{r}
dd4 <- LGBTQ %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicD <- ggplot(dd4,aes(x=infected,y=count,fill=treatment))
basicD + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicDD <- ggplot(dd4,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicDD + geom_bar(stat="identity", position = "dodge")
```

Studying the above generated graph provides a stark contrast  between LGBTQ infection data as compared to the earlier male and female results.  It seems that that while the majority of either LGBTQ group (both those who received the vaccine and the placebo) remained healthy, it seems that, of those who got COVID-19, an astounding ~85% of those cases had gotten the vaccine, compared to a meager 15% of placebo recipients.  While additional testing is necessary to corroborate these graphs, an early assessment suggests that the COVID-19 vaccine does NOT provide protection for members of the LGBTQ community.  


#### Numerical Results
```{r}
table3 <- xtabs(~treatment + infected, data=LGBTQ)
rowPerc(table3)
colPerc(table3)
```
The numerical results created, illustrated via the above table, further the messaging of the prior graphs generated in relation to the LGBTQ persons studied within the trial.  Although about half of the LGBTQ participants were not infected with COVID-19 at all, a larger proportion (3.43%) were reported to have taken the vaccine.  Of the people who WERE infected with COVID-19, though, a considerable 87% had received the vaccine; this further progresses the possibility that the vaccine offers no considerable immunity to members of the LGBTQ community, but further inferential testing is required to validate this.  

### Inferential Results
```{r}
chisq.test(table3)
chisqtestGC(table3)
fisher.test(table3)
LGBTQfisher <- fisher.test(table3)
```
Running the necessary inferential tests confirm our previously held assumptions about COVIVD-19 and vaccine efficacy as it relates to the LGBTQ members of this study.  Our observed counts show that 52 LGBTQ members who took the vaccine ended up getting COVID-19, compared to a relatively minuscule 8 members who had taken the placebo.  Our p-values for the Chi-Square (1e-08) and FisherExact (2e-09) are so little they're essentially zero; such magnitude itself points to a degree of statistical significance in our data.  Finally, the odds ratio was calculated to be .15, which can be understood to assert that for every 1 LGBTQ person who receives the vaccine and gets COVID-19, there are .15 odds of the same happening to an LGBTQ person getting the placebo and contracting COVID-19.  As such, it seems that the vaccine provides no efficacious protection against COVID-19 for members of the LGBTQ community.  


### Efficacy of Vaccine for LGBTQ Group
```{r}
LGBTQEfficacy <- LGBTQfisher$estimate/(1 + LGBTQfisher$estimate) * 100
LGBTQfisherlow <- LGBTQfisher$conf.int[1]
LGBTQfisherhigh <- LGBTQfisher$conf.int[2]
EfficacyLGBTQLow <- ((LGBTQfisherlow)/(1 + LGBTQfisherlow) * 100)
EfficacyLGBTQHigh <- ((LGBTQfisherhigh)/(1 + LGBTQfisherhigh) * 100)
```
The estimated efficacy of the vaccine among male patients is `r LGBTQEfficacy`%.  The 95% confidence interval for the efficacy of the vaccine pertaining to LGBTQ members is `r EfficacyLGBTQLow`% to `r EfficacyLGBTQHigh`%.




## Druggies

### Descriptive Results


#### Graphical Results
```{r}
dd5 <- DrugUsers %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicE <- ggplot(dd5,aes(x=infected,y=count,fill=treatment))
basicE + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicEE <- ggplot(dd5,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicEE + geom_bar(stat="identity", position = "dodge")
```

In running our graphical analysis of the statistics concerning treatment type and COVID-19 infection status among Drug Users in the study, our results seem to align closely with the outputs generated for LGBTQ participants.  While the essential majority of both groups of drug users seemed not to contract the disease, approximately 84% of those who were infected seemed to have received the vaccine, as compared to 16% of placebo recipients.  Without numerical and inferential testing, no final conclusions can be made, but it appears that, as with the case of the LGBTQ contributors, this COVID-19 vaccine does not readily provide immunity to persons who may be classified as Drug Users.  


#### Numerical Results
```{r}
table4 <- xtabs(~treatment + infected, data=DrugUsers)
rowPerc(table4)
colPerc(table4)
```
Our numerical results fortify the earlier GGPlot graphs created in that they're numerically corroborative.  For Drug Users who either got the placebo or vaccine, a majority remained health (99.4% and 96.7%, respectively).  However, our second table indicates that 85% of those who became COVID-19 positive had taken the vaccine; this pales in comparison to the 15% who had received the placebo.  Our tabular outputs, though they require further backing from our Chi-Square and Fisher Exact results, seem to also direct towards a lack of efficacy on the part of this COVID-19 vaccine, particularly among Drug Users.  

### Inferential Results
```{r}
chisq.test(table4)
chisqtestGC(table4)
fisher.test(table4)
DrugUserFisher <- fisher.test(table4)
```

As with the previous inferential tests, our Chi-Square and FisherExact results confer a sense of legitimacy about our data.  When we examine our observed counts in the Chi-Square test, we can see that 9 only Drug Users who contracted COVID-19 had received the placebo, compared to an astounding 51 users who were administered the vaccine.  These numbers rationalize the graphical and tabular data we explored earlier.  To add, our p-values for both the Chi-Square and Fisher Exact tests are essentially zero, so we can approach our data with a confident sense of statistical significance.  Our odds ratio was computed to be .18, close to our odds ratio in our inferential testing for LGBTQ members of this study (.15).  This statistic can simply be interpreted to say for every Drug User that contracts COVID-19 while having taken the vaccine, the odds of a Drug User contracting COVID-19 while having received a placebo are only .18 of the former.  As such, we can claim that our data/results are valid in their statements, and can conclude that the COVID-19 vaccine does not offer excellent immunity against COVID-19 for Drug Users.  


### Efficacy of Vaccine for Druggies 
```{r}
DrugUserEfficacy <- DrugUserFisher$estimate/(1 + DrugUserFisher$estimate) * 100
DrugUserfisherlow <- DrugUserFisher$conf.int[1]
DrugUserfisherhigh <- DrugUserFisher$conf.int[2]
EfficacyDrugUserLow <- ((DrugUserfisherlow)/(1 + DrugUserfisherlow) * 100)
EfficacyDrugUserHigh <- ((DrugUserfisherhigh)/(1 + DrugUserfisherhigh) * 100)
```
The estimated efficacy of the vaccine among male patients is `r DrugUserEfficacy`%.  The 95% confidence interval for the efficacy of the vaccine pertaining to drug users is `r EfficacyDrugUserLow`% to `r EfficacyDrugUserHigh`%.

# Conclusion
By amassing and analyzing a wealth of graphical, numerical, and inferential results, we can make an informed judgement on the efficacy of the COVID-19 vaccine in question, among males, females, LGBTQ people, and Drug Users.  Comparing our data among these groups shows great variation in whether or not the vaccine was effective in providing immunity.  We define an effective vaccine as one that provides 50% or more efficacy.  For males, our efficacy was estimated to be at around 67.66%, with a 95% confidence interval ranging from 62.51% to 72.51%.  Such a finding, along with our statistically significant results (see "Males") proves that the vaccine is indeed efficacious in preventing COVID-19 among males.  Such an argument can also be made for the females in the study; the efficacy estimate was computed to be around 58.83%, along with a 95% confidence interval from 53.5% to 64.01%.  It can be argued that, while the vaccine is less efficacious for women, there remains a degree of protection, and thus can still be recommended for defense against COVID-19.

However, cumulatively examining our data for the LGBTQ and Drug-Using participants in our study suggests the vaccine may not have been so valuable across all groups.  Our descriptive and inferential results for LGBTQ members of the study pointed towards little to no benefit with taking the vaccine against the placebo, and our efficacy data reflected this: the efficacy estimation among LGBTQ participants was around 12.78%, with a 95% confidence interval range of around 5.65% to 23.78%.  This is relatively poor compared to our efficacy yields among the males and females whose treatment options and infection statuses were recorded.  A similar outlook prevails for the observations relating to Drug Users, as their efficacy estimate was recorded at about 15.07%, along with a 95% confidence interval running from 7.11% to 26.78%.  Along with the latter data on LGBTQ members of the study, it's clear that the vaccine doesn't offer clear, substantial protection against COVID-19 for Drug Users, either.  

# Discussion
If we were to advise on the authorization or administration of this vaccine, our data has made it abundantly clear that it doesn't offer tangible results to every population group.  Inoculations may be more valuable for males and females, since they might provide up to 72.51% efficacy (for the former, with 64.01% for the latter).  However, a board contemplating whether or not to authorize this COVID-19 vaccine for the masses ought to reconsider promoting it to members of the LGBTQ or Drug-Using community, since efficacy levels are so low (12.78% and 15.07%, respectively) treatment with this product might be considered redundant.  To engineer a vaccine that better protects ALL demographics from COVID-19, the creators of this treatment should reexamine the ingredients and processes employed, and begin work again.  


