---
title: "Covid91 vaccine study Final2022"
author: "Rabi Elkhoury"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final_Data.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
options(digits = 1)
```

# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

```{r}
Males <- filter(FinalData, sex == "M")
Females <- filter(FinalData, sex == "F")
LGBTQ <- filter(FinalData, LGBTQ == "gay")
DrugUsers <- filter(FinalData, DrugUser == "yes")
```



## Males


### Descriptive Results

#### Graphical Results
```{r}
dd2 <- Males %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

Generating a barplot via ggplot for the data involving the incidence of MALES with COVID-19, along with whether or not they were treated with the vaccine, shows that, though a higher frequency of persons didn't get COVID-19 at all, most of those who did had taken the placebo.  Around 67% of males who contracted COVID-19 had taken the placebo, compared to only about 33% of males who had gotten the disease.  However, more testing is required to ensure such results are statistically significant.  


#### Numerical Results
```{r}
table1 <- xtabs(~treatment + infected, data=Males)
rowPerc(table1)
colPerc(table1)
```
Our numerical results corroborate the graphical descriptive results created above; we can see that, among Males who got COVID-19, 67.78% of them had taken the placebo, as compared to 32.22% of males who had gotten the vaccine.  While a majority of males didn't get COVID-19 at all, 2.63% of those men who received the placebo got the virus, as compared to only 1.28% of males who got the vaccine.  This further supports the notion that the vaccine aids in protection against the vaccine, but inferential results are required. 



### Inferential Results
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
malefisher <- fisher.test(table1)
malefisher
```
Multiple data points generated from our Chi-Square and Fisher Exact tests lend credibilty to our earlier tests.  The observed counts from the Chi-Square test show convey that over 100 more males who got the placebo ended up getting COVID-19, as compared to their vaccinated fellows.  Our odds ratio is 2.09, implying that males who get placebo-based treatment are 2.09 times as likely to get COVID-19 as compared to their vaccinated counterparts.  Furthermore, our p-value is infinitesimal (4.812e-11); such a small value suggests our results are statistically significant.  





### Efficacy of Vaccine for Males
```{r}
MaleEfficacy <- malefisher$estimate/(1 + malefisher$estimate) * 100
malefisherlow <- malefisher$conf.int[1]
malefisherhigh <- malefisher$conf.int[2]
EfficacyMalesLow <- ((malefisherlow)/(1 + malefisherlow) * 100)
EfficacyMalesHigh <- ((malefisherhigh)/(1 + malefisherhigh) * 100)
```
The estimated efficacy of the vaccine among male patients is `r MaleEfficacy`%.  The 95% confidence interval for the efficacy of the vaccine pertaining to males is `r EfficacyMalesLow`% to `r EfficacyMalesHigh`%.




## Females




### Descriptive Results





#### Graphical Results







#### Numerical Results



### Inferential Results




### Efficacy of Vaccine for Males




## LGBTQ

### Descriptive Results


#### Graphical Results




#### Numerical Results



### Inferential Results




### Efficacy of Vaccine for Males





## Druggies

### Descriptive Results


#### Graphical Results




#### Numerical Results



### Inferential Results




### Efficacy of Vaccine for Males






